Title

### 2018-09-22 thesis/paper/paper.tex

To improve the data quality, to help editors avoid common mistakes, and to clarify the usage of properties, Wikidata allows constraints to be defined on properties. They are, however, not firm rules, and exceptions can and should be made if necessary. A property constraint says something about the way a property may or may not be used. This can mean simple restrictions on the statement a property is part of, e.g. that only certain values are allowed, or more complex requirements on connected parts of the data, e.g. that the value of a statement with this property should have certain statements. Table \ref{tab_property_constraints} contains the existing property constraints as of September 2018. The constraints will be explained in detail and their modelling given. For constraint that cannot be modelled, reasoning for this assessment is included.

@@

Table \ref{tab_property_constraints} shows all constraints that are used as of September 2018. It is roughly sorted by how difficult it is to model the given constraint using logical rules, based on the required expressive features and pre-processing. These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset. Four constraints do not need any pre-processing except for loading the data into datalog as facts. Six can be solved if constant inequality is defined. This can be easily solved with brute force but will be improved in Section \ref{sec_constraints_using_inequality}. A further seven require negation of a specific fact which can be achieved with reasonable effort. The last ten are not deemed translatable because they require a feature that cannot be reasonably emulated in datalog. The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact. Other features needed for these constraints are value comparison, web requests to test if a link is valid, and regular expressions, which could theoretically be translated, but would require large amounts of pre-processing.

\begin{table}[H]
\caption{property constraints}\label{tab_property_constraints}
\begin{tabularx}{\textheight}{rlll}
\makecell{\# constrained\\ properties} & Identifier & Name & Required Features \\
\hline
1410 & Q53869507 & Scope &  \\
781 & Q21502838 & Conflicts with &  \\
27 & Q52004125 & Allowed entity types &  \\
17 & Q52558054 & None of &  \\
\hline
2843 & Q21502410 & Distinct value & Inequality \\
343 & Q21514353 & Allowed units & Inequality \\
303 & Q21510851 & Allowed qualifiers & Inequality \\
116 & Q21510859 & One of & Inequality \\
3 & Q52060874 & Single best value & Inequality \\
1 & Q52712340 & One-of qualifier value & Inequality \\
\hline
4265 & Q21503247 & Item requires statement & Negation (statement non-existence) \\
244 & Q21510864 & Value requires statement & Negation (statement non-existence) \\
119 & Q21510855 & Inverse & Negation (statement non-existence) \\
33 & Q21510862 & Symmetric & Negation (statement non-existence) \\
13 & Q21510857 & Multi-value & Negation, Count (up to two) \\
2909 & Q19474404 & Single value & Inequality, Negation (qualifier non-existence) \\
203 & Q21510856 & Mandatory qualifier & Negation (qualifier non-existence) \\
\hline
2613 & Q21503250 & Type & Negation (derived fact negation) \\
707 & Q21510865 & Value type & Negation (derived fact negation) \\
3021 & Q21502404 & Format & Regular expressions \\
188 & Q21510860 & Range & Value comparison \\
106 & Q52848401 & Integer & Regular expressions \\
65 & Q51723761 & No bounds & Negation \\
53 & Q21510852 & Commons link & Web request \\
10 & Q54554025 & Citation needed & Negation \\
6 & Q21510854 & Difference within range & Value comparison \\
1 & Q25796498 & Contemporary constraint & Negation \\
\end{tabularx}
\end{table}

@@

\subsection{Representation of Wikidata as Facts}\label{subsec_representation}
To represent the content of Wikidata as facts, the following predicates are introduced:
\begin{table}[H]
\caption{\textsc{edb} predicates}\label{tab_predicates}
\begin{tabular}{llp{8.5cm}}
Predicate & Positions & Description \\
\hline
\textsc{statement\_edb} & \makecell{statement ID \\ entity ID \\ property ID \\ value (entity ID or string)} & represents basic statements \\
\hline
\textsc{qualifier\_edb} & \makecell{statement ID \\ property ID \\ value} & represents qualifiers on statements \\
\hline
\textsc{reference\_edb} & \makecell{statement ID \\ reference hash \\ property ID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
\hline
\textsc{item\_edb} & entity ID & marks an entity as an item \\
\textsc{property\_edb} & entity ID & marks an entity as a property \\
\hline
\textsc{unit\_edb} & \makecell{value \\ unit ID} & denotes which unit a value has \\
\hline
\textsc{rank\_edb} & \makecell{statement ID \\ rank} & denotes the rank a statement has
\end{tabular}
\end{table}

@@

{\bf Constraint:} This subsection starts with a table showing basic properties of the constraint. These include the required data, relevant qualifiers, and possible violations of the constraint. The required data states which part of the data in relation to the constrained statement are restricted or need to exist. The qualifiers list their cardinality, meaning if they are required and if they may be used multiple times, and a short description. The violations list the conflicting (part of a) statement in relation to the constrained statement. The first constraint, \emph{property scope}, will be used as an example to explain the table in detail.

{\bf Rules:} Most constraints require more data than just the statements, qualifiers, and references. This can include additional data from Wikidata to be provided as facts using the introduced predicates (see Table \ref{tab_predicates}), inequalities to be established as a pre-processing step (see Section \ref{sec_constraints_using_inequality}), or preparations to determine specific kinds of negation (see Section \ref{sec_constraints_using_statement_non}). Afterwards, the rules necessary to determine violations are described.

@@

\subsection{Property scope (Q53869507)}
\paragraph{Constraint}
The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier, or reference.
\begin{table}[H]
\caption{property scope constraint (Q53869507)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statement, qualifier and reference \\
\hline
Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449), \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers, or references} \\
\hline
Violation & constrained property in a position that is not allowed \\
\hline
\end{tabularx}
\end{table}

@@

\subsection{Conflicts with (Q21502838)}
\paragraph{Constraint}
The \emph{conflicts with} constraint forbids statements on constrained entities whose predicate is a specified \emph{conflicting property}. If \emph{conflicting values} have been specified as well, the constraint is weakened and only violated by statements if in addition to the predicate being a \emph{conflicting property}, the object is one of the \emph{conflicting values}.
\begin{table}[H]
\caption{Conflicts-with constraint (Q21502838)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & entity statements \\
\hline
Qualifiers & \makecell{\emph{conflicting property} (P2306) -- 1 \\ property that may not be predicate of entity statements \\
\emph{conflicting values} (P2305) -- 0..* \\ values that may not be objects of entity statements with the \emph{conflicting property}} \\
\hline
Violation & \makecell{entity statement with the conflicting property \\ entity statement with the conflicting property and one of the conflicting values} \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{Allowed entity types constraint (Q52004125)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements and entity types\\
\hline
Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218), \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
\hline
Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{None of constraint (Q52558054)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements, qualifiers and references \\
\hline
Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier, or reference} \\
\hline
Violation & \makecell{constrained statement, qualifier, or reference with a forbidden value} \\
\hline
\end{tabularx}
\end{table}

@@

\subsubsection{Encoding constant strings to establish inequality}\label{subsubsec_encoding}
One way of determining inequality would be to take advantage of the fact that inequality is established in a set of unequal constants \(U\), which can be represented as a string. The idea is to determine the set of characters used in \(U\), set these characters as unequal as outlined in Section \ref{subsubsec_encoding} and associate each unequal constant with their characters.\\

An unequal constant \(u \in U\) can be represented a string and be written as \(c_0 \ldots c_{l - 1}\) with \(l = len(u)\) being the length of the string. Denote the set \(C\) as the set of characters occurring in unequal strings. Then \(l_{\max} = \max\limits_{\forall u \in U}(len(u))\) is the length of the longest entity string that occurs in the given data.
Based on this a group of predicates letter\(_i\)\_edb[8] is introduced where \(0 \equiv i \mod 7\), \(0 \leq i < l_{max} \).

For each unequal constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod 7, 0 \leq i < len(u)\) an atom using predicate \emph{letter\(_i\)\_edb} is created as 
\begin{align}
\begin{split}
\emph{letter}_i\emph{\_edb}(e, f(e, i), f(e, i+1), \ldots, f(e, i+6)
\end{split}
\begin{split}
f(e, i) = 
\begin{cases}
      c_i & i < len(e) \\
      \emph{none} & i \geq len(e)
\end{cases}
\end{split}
\end{align}

Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates \emph{letter\(_i\)\_edb} (with \emph{letters} short for \(\emph{letter}_i\emph{\_edb}(X, A_i, \ldots, A_{i+6}) \wedge\emph{letter}_i\emph{\_edb}(Y, B_i, \ldots, B_{i+6})\), \(i \leq j < i + 7\)):
\begin{align}
\textsc{unequal}(X, Y) \leftarrow \emph{letters} \wedge \textsc{unequal}(A_j, B_j)
\end{align}
Note that the facts fulfilling \(\textsc{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all unequal constants \(U\).

@@

\subsection{Distinct values (Q21502410)}\label{subsec_distinct_values}
\paragraph{Constraint}
The \emph{distinct values} constraint forbids the existence of two constrained statements with the same object.
\begin{table}[H]
\caption{distinct values constraint (Q21502410)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements \\
\hline
Violations & two constrained statements with same value \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}
The constraint can be expanded to cover qualifiers and references as well by setting a \emph{constraint scope} (P4680) qualifier with the respective value. However, as of September 2018 there is no constraint making use of this.

@@

\begin{table}[H]
\caption{Allowed units constraint (Q21514353)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements \\
\hline
Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements (\emph{none} for dimensionless units)} \\
\hline
Violation & \makecell{constrained statement with a unit unequal to all allowed units} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{Allowed qualifiers constraint (Q21510851)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & qualifiers on constrained statements \\
\hline
Qualifiers & \makecell{\emph{allowed qualifier} (P2306) -- 1..* \\ properties that may appear in predicates of qualifiers on constrained statements} \\
\hline
Violation & \makecell{constrained statement with a qualifier unequal to all allowed qualifiers} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{One-of constraint (Q21510859)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements, qualifiers, and references \\
\hline
Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
\hline
Violation & \makecell{constrained statement with a value unequal to all allowed values} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The one-of constraint fits the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers, and references. Inequality is established between all \emph{allowed values} and the values of statements, qualifiers, and references, respectively. With \(A\) the set of allowed values the rules for triple, qualifier, and reference are:
\begin{align}
\begin{split}
\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
\end{split}\\
\begin{split}
\textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, V) \leftarrow \\
&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
\end{split}\\
\begin{split}
\textsc{violation\_reference}&(S, \emph{constrainedProperty}, V) \leftarrow \\
&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V) {\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
\end{split}
\end{align}

@@

\begin{table}[H]
\caption{One-of qualifier value constraint (Q52712340)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements and their qualifiers \\
\hline
Qualifiers: & \makecell{\emph{qualifier property} (P2306) -- 1 \\ property that, if used as predicate, constrains a qualifier \\
\emph{qualifier values} (P2305) -- 1..* \\ values that are allowed for qualifiers on constrained statements \\ with the \emph{qualifier property} as predicate}\\
\hline
Violation & \makecell{constrained statement with a qualifier with the qualifier property \\ and a value unequal to all qualifier values} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\section{Constraints using Statement non-Existence}\label{sec_constraints_using_statement_non}
Several constraints are violated if specific statements do not exist in relation to constrained statements. It is not possible to simply negate said statements and pre-processing every required statement would defeat the purpose of implementing them in datalog. However, non-existence can be derived utilising the order of statements. Three predicates are introduced: \textsc{first}with arity 1, \textsc{last}with arity 2, and \textsc{next}with arity 2, which denote the first and last statement of an item and the following statement to another using their statement IDs.\\

@@

\subsection{Item requires statement (Q21503247)}\label{subsec_item_requires_statement}
\paragraph{Constraint}
The \emph{item requires statement} constraint requires an additional statement on an item with a statement using the constrained property. It can be even more restrictive by also requiring the value of said additional statement to be in a set of \emph{allowed values}.
\begin{table}[H]
\caption{item requires statement constraint (Q21503247)}\label{tab_item_requires_statements}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & item statements \\
\hline
Qualifiers & \makecell{\emph{required property} (P2306) -- 1 \\ property that must be predicate of an item statement \\
\emph{allowed value} (P2305) -- 0..* \\ value that must be object of the item statement} \\
\hline
Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
\hline
Rule features & negation \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{Multi-value constraint (Q21510857)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & item statements \\
\hline
Violation & \makecell{constrained item with exactly one constrained statement} \\
\hline
Dependency & all item statements \\
\hline
Rule features & count \\
\hline
\end{tabularx}
\end{table}

@@

\subsection{Single value (Q19474404)}\label{subsec_single_value}
\paragraph{Constraint}
The \emph{single value} constraint states that the constrained property may only be used once per item. A constrained statement marked with the single value constraint can have \emph{separators}, which define exceptions under specific circumstances.
\begin{table}[H]
\caption{single value constraint (Q19474404)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements and their qualifiers, constrained qualifiers and references \\
\hline
Qualifiers & \makecell{\emph{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception} \\
\hline
Violation & \makecell{two constrained statements with same item and value \\ (exception: statements with separators as qualifiers with different values)} \\
\hline
Rule features & inequality, partially negation \\
\hline
\end{tabularx}
\end{table}

@@

At the beginning the content of the Wikidata JSON-Dump, read via WDKT, is transformed into facts readable by VLog4J. These take the form of comma-separated value files, one for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram.

At the same time the Inequaliy Helper, receiving the Facts from WDTK, creates all files necessary to establish inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen. These files are constraint type specific, whereas the "Facts" are not.

>>>>

To improve the data quality, to help editors avoid common mistakes, and to clarify the usage of properties, Wikidata allows constraints to be defined on properties. They are, however, not firm rules, and exceptions can and should be made if necessary. A property constraint says something about the way a property may or may not be used. This can mean simple restrictions on the statement a property is part of, e.g. that only certain values are allowed, or more complex requirements on connected parts of the data, e.g. that the value of a statement with this property should have certain statements. Table \ref{tab_property_constraints} contains the existing constraint types as of September 2018. The constraint types will be explained in detail and their modelling given. For constraint that cannot be modelled, reasoning for this assessment is included.

@@

Table \ref{tab_property_constraints} shows all constraints that are used as of September 2018. It contains the name of the constraint and its Q-ID, followed by the number of constraint declarations and the features required to model the constraint type. The table is roughly sorted by how difficult it is to model the given constraint using logical rules, based on the required expressive features and pre-processing. The horizontal lines group  These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset. Four constraints do not need any pre-processing except for loading the data into datalog as facts. Six can be solved if constant inequality is defined. This can be easily solved with brute force but will be improved in Section \ref{sec_constraints_using_inequality}. A further seven require negation of a specific fact which can be achieved with reasonable effort. The last ten are not deemed translatable because they require a feature that cannot be reasonably emulated in datalog. The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact. Other features needed for these constraints are value comparison, web requests to test if a link is valid, and regular expressions, which could theoretically be translated, but would require large amounts of pre-processing.

\begin{table}[H]
\caption{property constraints}\label{tab_property_constraints}
\begin{tabular}{llrp{6.5cm}}
Name & Identifier & \makecell{\# constrained \\ properties} & Required Features \\
\hline
Scope & Q53869507 & 1410 & \\
Conflicts with & Q21502838 & 781 & \\
Allowed entity types & Q52004125 & 27 & \\
None of & Q52558054 & 17 & \\
\hline
Distinct value & Q21502410 & 2843 & Inequality \\
Allowed units & Q21514353 & 343 & Inequality \\
Allowed qualifiers & Q21510851 & 303 & Inequality \\
One of & Q21510859 & 116 & Inequality \\
Single best value & Q52060874 & 3 & Inequality \\
One-of qualifier value & Q52712340 & 1 & Inequality \\
\hline
Item requires statement & Q21503247 & 4265 & Negation (statement non-existence) \\
Value requires statement & Q21510864 & 244 & Negation (statement non-existence) \\
Inverse & Q21510855 & 119 & Negation (statement non-existence) \\
Symmetric & Q21510862 & 33 & Negation (statement non-existence) \\
Multi-value & Q21510857 & 13 & Negation, Count (up to two)  \\
Single value & Q19474404 & 2909 & Inequality, Negation (qualifier non-existence) \\
Mandatory qualifier & Q21510856 & 203 & Negation (qualifier non-existence) \\
\hline
Type & Q21503250 & 2613 & Negation (derived fact negation) \\
Value type & Q21510865 & 707 & Negation (derived fact negation) \\
Format & Q21502404 & 3021 & Regular expressions \\
Range & Q21510860 & 188 & Value comparison \\
Integer & Q52848401 & 106 & Regular expressions \\
No bounds & Q51723761 & 65 & Negation  \\
Commons link & Q21510852 & 53 & Web request \\
Citation needed & Q54554025 & 10 & Negation \\
Difference within range & Q21510854 & 6 & Value comparison \\
Contemporary constraint & Q25796498 & 1 & Negation  \\
\end{tabular}
\end{table}

@@

\subsection{Representation of Wikidata as Facts}\label{subsec_representation}
To represent the content of Wikidata as facts, the following predicates are introduced:
\begin{table}[H]
\caption{\textsc{edb} predicates}\label{tab_predicates}
\begin{tabular}{llp{8cm}}
Predicate & Positions & Description \\
\hline
\textsc{statement\_edb} & \makecell{statement ID \\ entity ID \\ property ID \\ value (entity ID or string)} & represents basic statements \\
\hline
\textsc{qualifier\_edb} & \makecell{statement ID \\ property ID \\ value} & represents qualifiers on statements \\
\hline
\textsc{reference\_edb} & \makecell{statement ID \\ reference hash \\ property ID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
\hline
\textsc{item\_edb} & entity ID & marks an entity as an item \\
\textsc{property\_edb} & entity ID & marks an entity as a property \\
\hline
\textsc{unit\_edb} & \makecell{value \\ unit ID} & denotes which unit a value has \\
\hline
\textsc{rank\_edb} & \makecell{statement ID \\ rank} & denotes the rank a statement has
\end{tabular}
\end{table}

@@

{\bf Constraint:} This subsection starts with a table showing basic properties of the constraint. These include the required data, relevant qualifiers, and possible violations of the constraint. The required data states which facts need to be loaded. The qualifiers list their cardinality, meaning if they are required and if they may be used multiple times, and a short description. The violations list the conflicting (part of a) statement in relation to the constrained statement. The first constraint, \emph{property scope}, will be used as an example to explain the table in detail.

{\bf Rules:} Most constraints require more data than just the statements, qualifiers, and references. This can include inequalities to be established as a pre-processing step (see Section \ref{sec_constraints_using_inequality}) or preparations to determine specific kinds of negation (see Section \ref{sec_constraints_using_statement_non}). Afterwards, the rules necessary to determine violations are described.

@@

\subsection{Property scope (Q53869507)}
\paragraph{Constraint}
The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier, or reference.
\begin{table}[H]
\caption{property scope constraint (Q53869507)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & statements, qualifiers, and references\\
\hline
Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449), \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers, or references} \\
\hline
Violation & constrained property in a position that is not allowed \\
\hline
\end{tabularx}
\end{table}

@@

\subsection{Conflicts with (Q21502838)}
\paragraph{Constraint}
The \emph{conflicts with} constraint forbids statements on constrained entities whose predicate is a specified \emph{conflicting property}. If \emph{conflicting values} have been specified as well, the constraint is weakened and only violated by statements if in addition to the predicate being a \emph{conflicting property}, the object is one of the \emph{conflicting values}.
\begin{table}[H]
\caption{Conflicts-with constraint (Q21502838)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & statements \\
\hline
Qualifiers & \makecell{\emph{conflicting property} (P2306) -- 1 \\ property that may not be predicate of entity statements \\
\emph{conflicting values} (P2305) -- 0..* \\ values that may not be objects of entity statements with the \emph{conflicting property}} \\
\hline
Violation & \makecell{entity statement with the conflicting property \\ entity statement with the conflicting property and one of the conflicting values} \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{Allowed entity types constraint (Q52004125)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & statements, entity type (item or property)\\
\hline
Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218), \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
\hline
Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{None of constraint (Q52558054)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & statements, qualifiers, and references \\
\hline
Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier, or reference} \\
\hline
Violation & \makecell{constrained statement, qualifier, or reference with a forbidden value} \\
\hline
\end{tabularx}
\end{table}

@@

\subsubsection{Encoding constant strings to establish inequality}\label{subsubsec_encoding}
One way of determining inequality would be to take advantage of the fact that inequality is established in a set of constants \(U\), which can be represented as a string. The idea is to determine the set of characters used in \(U\), set these characters as unequal as outlined in Section \ref{subsubsec_encoding} and associate each unequal constant with their characters. One possibility would be to create one predicate to associate all characters of a string with said string. This predicates arity would likely be very large and thus not efficient to handle. If each position was described with its own binary predicate, this could be avoided. It would however require the string to be saved to disk repeatedly, once for each position in the string.\\

The approach chosen here thus splits the constant string into chunks of length \(c\) and saves these with one predicate per chunk. Varying the chunk length allows balancing the disk space requirements with potential performance increases gained through shorter predicates.

A constant \(u \in U\) can be represented a string and be written as \(c_0 \ldots c_{l - 1}\) with \(l = len(u)\) being the length of the string. Denote the set \(C\) as the set of characters occurring in strings. Then \(l_{\max} = \max\limits_{\forall u \in U}(len(u))\) is the length of the longest entity string that occurs in the given data.
Based on this a group of predicates letter\(_i\)\_edb[c+1] is introduced where \(0 \equiv i \mod c\), \(0 \leq i < l_{max} \).

For each constant string \(u = c_0 \ldots c_{l - 1}\) and \(0 \equiv i \mod c, 0 \leq i < len(u)\) an atom using predicate \emph{letter\(_i\)\_edb} is created as 
\begin{align}
\begin{split}
\emph{letter}_i\emph{\_edb}(e, f(e, i), f(e, i+1), \ldots, f(e, i+c-1)
\end{split}
\begin{split}
f(e, i) = 
\begin{cases}
      c_i & i < len(e) \\
      \emph{none} & i \geq len(e)
\end{cases}
\end{split}
\end{align}

Now that inequality between all characters and the link between entities and their characters are introduced, the inequality between two entities can be derived by creating the following rules for all predicates \emph{letter\(_i\)\_edb} (with \emph{letters} short for \(\emph{letter}_i\emph{\_edb}(X, A_i, \ldots, A_{i+6}) \wedge\emph{letter}_i\emph{\_edb}(Y, B_i, \ldots, B_{i+c-1})\), \(i \leq j < i + c\)):
\begin{align}
\textsc{unequal}(X, Y) \leftarrow \emph{letters} \wedge \textsc{unequal}(A_j, B_j)
\end{align}
Note that the facts fulfilling \(\textsc{unequal}(A_j, B_j)\) are created by using the approach outlined in Section \ref{subsubsec_naive} for the set of characters \(C\) in all constants \(U\).

Another possibility to apply this approach would be to encode inequality for the string hashes instead of the strings themselves. The string and its hash would then be associated via another predicate. Since generating these hashes is no trivial task and the additional step from constant to hash would complicate the rules presented below, this was not done here.

@@

\subsection{Distinct values (Q21502410)}\label{subsec_distinct_values}
\paragraph{Constraint}
The \emph{distinct values} constraint forbids the existence of two constrained statements with the same object.
\begin{table}[H]
\caption{distinct values constraint (Q21502410)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements; \\
inequality for statement IDs} \\
\hline
Violations & two constrained statements with same value \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}
The constraint can be expanded to cover qualifiers and references as well by setting a \emph{constraint scope} (P4680) qualifier with the respective value. However, as of September 2018 there is no constraint making use of this.

@@

\begin{table}[H]
\caption{Allowed units constraint (Q21514353)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements, qualifiers, and references; \\
inequality for units} \\
\hline
Qualifiers & \makecell{\emph{allowed unit} (P2305) -- 1..* \\ units that are allowed on values of constrained statements \\ (\emph{none} for dimensionless units)} \\
\hline
Violation & \makecell{constrained statement with a unit unequal to all allowed units} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{Allowed qualifiers constraint (Q21510851)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements and qualifiers; \\
inequality for qualifier properties} \\
\hline
Qualifiers & \makecell{\emph{allowed qualifier} (P2306) -- 1..* \\ properties that may appear in predicates of qualifiers on constrained statements} \\
\hline
Violation & \makecell{constrained statement with a qualifier unequal to all allowed qualifiers} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{One-of constraint (Q21510859)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements, qualifiers, and references; \\
inequality for values of statements, qualifiers, and references} \\
\hline
Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
\hline
Violation & \makecell{constrained statement with a value unequal to all allowed values} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The one-of constraint fits the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers, and references. Inequality is established between all \emph{allowed values} and the values of statements, qualifiers, and references, respectively. With \(A\) the set of allowed values the rules for triple, qualifier, and reference are:
\begin{align}
\begin{split}
\textsc{violation\_statement}&(S, E, \emph{constrainedProperty}, V) \leftarrow \\
&\textsc{statement\_edb}(S, E, \emph{constrainedProperty}, V) \\
&{\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
\end{split}\\
\begin{split}
\textsc{violation\_qualifier}&(S, \emph{constrainedProperty}, V) \leftarrow \\
&\textsc{qualifier\_edb}(S, \emph{constrainedProperty}, V) \\
&{\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
\end{split}\\
\begin{split}
\textsc{violation\_reference}&(S, \emph{constrainedProperty}, V) \leftarrow \\
&\textsc{reference\_edb}(S, H, \emph{constrainedProperty}, V) \\
&{\textstyle\bigwedge}_{a \in A} \textsc{unequal}(a, V)
\end{split}
\end{align}

@@

\begin{table}[H]
\caption{One-of qualifier value constraint (Q52712340)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements and qualifiers; \\
inequality for values of qualifiers} \\
\hline
Qualifiers: & \makecell{\emph{qualifier property} (P2306) -- 1 \\ property that, if used as predicate, constrains a qualifier \\
\emph{qualifier values} (P2305) -- 1..* \\ values that are allowed for qualifiers on constrained statements \\ with the \emph{qualifier property} as predicate}\\
\hline
Violation & \makecell{constrained statement with a qualifier with the qualifier property \\ and a value unequal to all qualifier values} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\section{Constraints Using Statement Non-Existence}\label{sec_constraints_using_statement_non}
Several constraints are violated if specific statements do not exist in relation to constrained statements. It is not possible to simply negate said statements and pre-processing every required statement would defeat the purpose of implementing them in datalog. However, non-existence can be derived utilising the order of statements. Three predicates are introduced: \textsc{first}with arity 1, \textsc{last}with arity 2, and \textsc{next}with arity 2, which denote the first and last statement of an item and the following statement to another using their statement IDs.\\

@@

\subsection{Item requires statement (Q21503247)}\label{subsec_item_requires_statement}
\paragraph{Constraint}
The \emph{item requires statement} constraint requires an additional statement on an item with a statement using the constrained property. It can be even more restrictive by also requiring the value of said additional statement to be in a set of \emph{allowed values}.
\begin{table}[H]
\caption{item requires statement constraint (Q21503247)}\label{tab_item_requires_statements}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements; \\
inequality for statement properties and values; \\
statement order} \\
\hline
Qualifiers & \makecell{\emph{required property} (P2306) -- 1 \\ property that must be predicate of an item statement \\
\emph{allowed value} (P2305) -- 0..* \\ value that must be object of the item statement} \\
\hline
Violation & \makecell{no item statement with required property \\ all item statements with required property have no allowed value} \\
\hline
Rule features & negation \\
\hline
\end{tabularx}
\end{table}

@@

\begin{table}[H]
\caption{Multi-value constraint (Q21510857)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements; \\
inequality for statement properties; \\
statement order} \\
\hline
Violation & \makecell{constrained item with exactly one constrained statement} \\
\hline
Dependency & all item statements \\
\hline
Rule features & count \\
\hline
\end{tabularx}
\end{table}

@@

\subsection{Single value (Q19474404)}\label{subsec_single_value}
\paragraph{Constraint}
The \emph{single value} constraint states that the constrained property may only be used once per item. A constrained statement marked with the single value constraint can have \emph{separators}, which define exceptions under specific circumstances.
\begin{table}[H]
\caption{single value constraint (Q19474404)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & \makecell{statements, qualifiers, and references; \\
inequality for statement IDs and qualifier values; \\
qualifier order} \\
\hline
Qualifiers & \makecell{\emph{separator} (P4155) -- 0..* \\ qualifiers that, when used on constrained statements, can warrant an exception} \\
\hline
Violation & \makecell{two constrained statements with same item and value \\ (exception: statements with separators as qualifiers with different values)} \\
\hline
Rule features & inequality, partially negation \\
\hline
\end{tabularx}
\end{table}

@@

At the beginning the content of the Wikidata JSON-Dump, read via WDKT, is transformed into facts readable by VLog4J. These take the form of comma-separated value files, one for each predicate in Section \ref{subsec_representation}, denoted as "Facts" in the diagram. The statement and qualifier orders are generated as well. These consist of three files each for the first, next, and last predicate.

At the same time the Inequality Helper, receiving the Facts from WDTK, creates all files necessary to establish inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen. These files are constraint type specific, whereas the "Facts" are not.
